SubsidyExplorer - Step 1: Data Wrangling

### 2020-08-06 scripts/01_data_wrangling.Rmd
added new FAO data and RAM data

```{r}
# Packages
library(knitr) 
library(countrycode) # country name matching
library(here)
library(WDI) # World Bank GDP and population numbers by country
library(janitor) # cleaning column names
library(tidyverse)
library(readxl)

@@

# Source helper file
source(here::here("scripts", "00_helpers.R"))

@@

### Paths to fisheries subsidies data files
sumaila_subsidies_path <- here::here("data", "raw", "Sumaila-et-al-2019-subsidies", "Subsidies_PEW_Final.xlsx")
oecd_fse_path <- here::here("data", "raw", "OECD-2019-fse", "FISH_FSE_10032020183349581.csv")
schuhbauer_subsidies_ssf_path <- here::here("data", "raw", "Schuhbauer-et-al-2017-ssf-subsidies", "Subsidies_SSF_split_2018.csv")

### Paths to other fisheries data files
fao_production_path <- here::here("data", "raw", "FAO-2019-capture-production", "Capture_2019.1.0/")
#fao_value_path
melnychuk_price_path <- here::here("data", "raw", "Melnychuk-et-al-2013-price", "Melnychuk_Exvessel_Price_Database.csv")
ram_sadb_path <- here::here("data", "raw", "RAM-2018-stock-assessment-database", "RLSADB v4.40", "DB Files With Assessment Data", "RLSADB v4.40 (assessment data only).xlsx")
costello_upsides_path <- here::here("data", "raw", "Costello-et-al-2016-upsides", "ProjectionData_2012.csv")
melnychuk_fmi_path <- here::here("data", "raw", "Melnychuk-et-al-2017-fmi", "fmi_values_country.csv")
iuu_vessel_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")

### Paths to other necessary files
cap_tier_path <- here::here("data", "raw", "USA-Cap-Tier-Proposal", "US_cap_tier_data.csv")
country_lookup_path <- here::here("data", "lookup-tables", "country_dependencies.csv")
fao_regions_for_ram_lookup_path <- here::here("data", "lookup-tables", "fao_areas_for_ram_stocks_v4.491.csv")
```

@@

- Proposed tier assignments for a cap-based approach to limiting subsidies as included in the proposal made to the WTO by the US Delegation on May 3, 2019 [TN/RL/GEN/197/Rev.1].

@@

@@

## Updated estimates and analysis of global fisheries subsidies from Sumaila et al. (2019)

@@

# Load lookup table for classification scheme
sumaila_lookup <- read_csv(here::here("data", "lookup-tables", "sumaila_subsidies_classification_lookup_table.csv"))

@@

# Load OECD lookup table
oecd_fse_lookup <- read_csv(here::here("data", "lookup-tables", "oecd_fse_classification_lookup_table.csv")) %>%
  distinct(category, category_name, type, type_code, type_name)

@@

@@

# Put them back together
fao_production_dat_edit <- fao_production_dat %>%
  dplyr::filter(iso3 != "EU") %>%
  bind_rows(fao_production_dat_eu) %>%
  bind_rows(fao_production_dat_us) %>%
  mutate(variable = "capture_production",
         units = "tonnes",
         source = "FAO (2019)")

write_csv(fao_production_dat_edit, paste0(data_dir, "fao_2019_capture_production_tidy.csv"))

@@

write_csv(fao_production_dat_edit_groups, paste0(data_dir, "fao_2019_capture_production_isscaap_groups_tidy.csv"))
write_csv(fao_production_dat_edit_groups, paste0(app_data_dir, "fao_2019_capture_production_isscaap_groups_tidy.csv"))
```

@@

# CPI for 2012 - 2018 to convert USD
cpi_2012 <- 229.594
cpi_2013 <- 232.957
cpi_2014 <- 236.736
cpi_2015 <- 237.017
cpi_2016 <- 240.007
cpi_2017 <- 245.120

# Do extrapolation for missing years
melnychuk_price_dat_extrapolation <- melnychuk_price_dat %>%
  complete(nesting(scientific_name, asfis_species, isscaap_group, group_for_pairing), year = 2000:2017) %>%
  left_join(melnychuk_price_dat_mry, by = c("scientific_name", "asfis_species", "isscaap_group", "group_for_pairing")) %>%
  mutate(exvessel_adjusted = case_when(year <= 2012 ~ exvessel,
                                       year == 2013 ~ exvessel_for_extrap * (cpi_2013/cpi_2012),
                                       year == 2014 ~ exvessel_for_extrap * (cpi_2014/cpi_2012),
                                       year == 2015 ~ exvessel_for_extrap * (cpi_2015/cpi_2012),
                                       year == 2016 ~ exvessel_for_extrap * (cpi_2016/cpi_2012),
                                       year == 2017 ~ exvessel_for_extrap * (cpi_2017/cpi_2012))) %>%
  dplyr::select(-exvessel, -exvessel_for_extrap) %>%
  rename(exvessel = exvessel_adjusted)

@@

# Look at NAs - 9398 entries
combine_nas <- combine_catch_price_dat %>%
  dplyr::filter(is.na(exvessel_inclusive))

@@

# Calculate revenues and export
landed_value_dat <- combine_catch_price_dat %>%
  dplyr::select(-exvessel, -exvessel_group_avg, -exvessel_group_avg_allyears) %>%
  rename(exvessel = exvessel_inclusive) %>%
  mutate(landed_value = value*exvessel) %>%
  group_by(iso3, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(landed_value, na.rm = T)) %>%
  mutate(variable = "landed_value",
         units = paste(year, "US$"),
         source = "FAO (2019); Melnychuk et al. (2017)")

@@

## Stock status database
```{r}
# Load data
costello_upsides_dat <- read_csv(costello_upsides_path) %>%
  clean_names(case = "snake")

# We need to split the entries that span multiple FAO regions...  
costello_upsides_dat_edit <- costello_upsides_dat %>%
  mutate(region_fao_chr = as.character(region_fao),
         n_regions = nchar(region_fao_chr)/2,
         region_fao_split = str_split(gsub("(.{2})", "\\1 ", region_fao_chr), " ")) %>%
  unnest(cols = c(region_fao_split)) %>%
  mutate(region_fao_split = as.numeric(region_fao_split)) %>%
  dplyr::filter(region_fao_split != "") %>%
  mutate(iso3 = case_when(country == "Channel Islands" ~ "JEY",
                          country == "Micronesia" ~ "FSM",
                          country == "Zanzibar" ~ "TZA",
                          TRUE ~ countrycode(country, "country.name", "iso3c")))

@@

# Section 4 - Other Data

>>>>

```{r}
# Packages
library(knitr) # Necessary to convert file to .html
library(countrycode) # Country name matching
library(here) # File path handling
library(WDI) # World Bank World Development Indicators data
library(janitor) # Column name handling
library(tidyverse) # General data wrangling
library(readxl) # Load .xlsx files

@@

# Source file with helper functions (used throughout analysis)
source(here::here("scripts", "00_helpers.R"))

@@

### Paths to fisheries subsidies data files
sumaila_subsidies_path <- here::here("data", "raw", "Sumaila-et-al-2019-subsidies", "Subsidies_PEW_Final.xlsx")
oecd_fse_path <- here::here("data", "raw", "OECD-fisheries-support-estimates", "FISH_FSE_10032020183349581.csv")
schuhbauer_subsidies_ssf_path <- here::here("data", "raw", "Schuhbauer-et-al-2017-ssf-subsidies", "Subsidies_SSF_split_2018.csv")

### Paths to other fisheries data files
fao_production_path <- here::here("data", "raw", "FAO-global-capture-production", "Capture_2020.1.0/")
#fao_value_path
melnychuk_price_path <- here::here("data", "raw", "Melnychuk-et-al-2013-price", "Melnychuk_Exvessel_Price_Database.csv")
ram_sadb_path <- here::here("data", "raw", "RAM-legacy-stock-assessment-database", "RAMLDB v4.491", "DB Files With Assessment Data", "RAMCore/")
costello_upsides_path <- here::here("data", "raw", "Costello-et-al-2016-upsides", "ProjectionData_2012.csv")
melnychuk_fmi_path <- here::here("data", "raw", "Melnychuk-et-al-2017-fmi", "fmi_values_country.csv")
iuu_vessel_path <- here::here("data", "raw", "combined-IUU-vessel-list", "IUUList-20190430.csv")

### Paths to other necessary files
cap_tier_path <- here::here("data", "raw", "USA-cap-tier-proposal", "US_cap_tier_data.csv")
country_lookup_path <- here::here("data", "lookup-tables", "country_dependencies.csv")
fao_regions_for_ram_lookup_path <- here::here("data", "lookup-tables", "fao_areas_for_ram_stocks_v4.491.csv")
sumaila_lookup_path <- here::here("data", "lookup-tables", "sumaila_subsidies_classification_lookup_table.csv")
schuhbauer_lookup_path <- here::here("data", "lookup-tables", "schuhbauer_subsidies_classification_lookup_table.csv")
oecd_fse_lookup_path <- here::here("data", "lookup-tables", "oecd_fse_classification_lookup_table.csv")
```

@@

- Proposed tier assignments for a cap-based approach to limiting subsidies as included in the proposal made to the WTO by the delegations of Austraila and the USA on May 3, 2019 [TN/RL/GEN/197/Rev.1].

@@

- All subsidy types used by Sumaila et al. (2019)

- All subsidy types used by the OECD FSE database

@@

## Estimates of global fisheries subsidies from Sumaila et al. (2019)

@@

# Load lookup table for classification scheme
sumaila_lookup <- read_csv(sumaila_lookup_path)

@@

# Load OECD lookup table
oecd_fse_lookup <- read_csv(oecd_fse_lookup_path) %>%
  distinct(category, category_name, type, type_code, type_name)

@@

## Updated estimates of small-scale fisheries subsides from Schuhbauer et al. (2017)

Note: these data are unpublished updates (in 2018 US\$) of the data described in Schuhbauer et al. (2017).

```{r}
# Load and wrangle data (2018 US$)
schuhbauer_subsidies_ssf_dat <- read_csv(schuhbauer_subsidies_ssf_path) %>%
  clean_names(case = "snake") %>%
  mutate(iso3 = case_when(country == "Dominican Rp" ~ "DOM",
                          country == "Micronesia" ~ "FSM",
                          country == "Untd Arab Em" ~ "ARE",
                          TRUE ~ countrycode(country, "country.name", "iso3c"))) %>%
  mutate(prop_ssf = ssf_usd/constant_2018_usd) %>%
  dplyr::select(iso3, old_type_name = type, prop_ssf) %>%
  mutate(source = "Schuhbauer et al. (2017)") %>%
  arrange(iso3)
schuhbauer_subsidies_ssf_dat$prop_ssf[is.nan(schuhbauer_subsidies_ssf_dat$prop_ssf)] <- 0 # handling dividing by 0

# Load lookup table
schuhbauer_lookup <- read_csv(schuhbauer_lookup_path)

# Clean up the naming/spelling of the classification scheme as well as data type
schuhbauer_subsidies_ssf_dat_edit <- schuhbauer_subsidies_ssf_dat %>%
  left_join(schuhbauer_lookup, by = c("old_type_name")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, prop_ssf, source)

# Deal with the EU
schuhbauer_subsidies_ssf_dat_eu <- read_csv(schuhbauer_subsidies_ssf_path) %>%
  clean_names(case = "snake") %>%
  mutate(iso3 = case_when(country == "Dominican Rp" ~ "DOM",
                          country == "Micronesia" ~ "FSM",
                          country == "Untd Arab Em" ~ "ARE",
                          TRUE ~ countrycode(country, "country.name", "iso3c"))) %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, type) %>%
  summarize(constant_2018_usd = sum(constant_2018_usd, na.rm = T),
            ssf_usd = sum(ssf_usd, na.rm = T)) %>%
  ungroup() %>%
  mutate(prop_ssf = ssf_usd/constant_2018_usd) %>%
  dplyr::select(iso3 = entity, old_type_name = type, prop_ssf) %>%
  mutate(source = "Schuhbauer et al. (2017)") %>%
  left_join(schuhbauer_lookup, by = c("old_type_name")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, prop_ssf, source)

# Add the EU back in
schuhbauer_subsidies_ssf_dat_edit <- schuhbauer_subsidies_ssf_dat_edit %>%
  bind_rows(schuhbauer_subsidies_ssf_dat_eu)

# Save
write_csv(schuhbauer_subsidies_ssf_dat_edit, paste0(data_dir, "schuhbauer_ssf_subsidies_proportions.csv"))
write_csv(schuhbauer_subsidies_ssf_dat_edit, paste0(app_data_dir, "schuhbauer_ssf_subsidies_proportions.csv"))

@@

# Put them back together
fao_production_dat_edit <- fao_production_dat %>%
  dplyr::filter(iso3 != "EU") %>%
  bind_rows(fao_production_dat_eu) %>%
  bind_rows(fao_production_dat_us) %>%
  mutate(variable = "capture_production",
         units = "tonnes",
         source = "FAO (2020)")

write_csv(fao_production_dat_edit, paste0(data_dir, "fao_2020_capture_production_tidy.csv"))

@@

write_csv(fao_production_dat_edit_groups, paste0(data_dir, "fao_2020_capture_production_isscaap_groups_tidy.csv"))
write_csv(fao_production_dat_edit_groups, paste0(app_data_dir, "fao_2020_capture_production_isscaap_groups_tidy.csv"))
```

@@

# CPI for 2012 - 2018 to convert USD
cpi_2012 <- 229.594
cpi_2013 <- 232.957
cpi_2014 <- 236.736
cpi_2015 <- 237.017
cpi_2016 <- 240.007
cpi_2017 <- 245.120
cpi_2018 <- 251.107

# Do extrapolation for missing years
melnychuk_price_dat_extrapolation <- melnychuk_price_dat %>%
  complete(nesting(scientific_name, asfis_species, isscaap_group, group_for_pairing), year = 2000:2018) %>%
  left_join(melnychuk_price_dat_mry, by = c("scientific_name", "asfis_species", "isscaap_group", "group_for_pairing")) %>%
  mutate(exvessel_adjusted = case_when(year <= 2012 ~ exvessel,
                                       year == 2013 ~ exvessel_for_extrap * (cpi_2013/cpi_2012),
                                       year == 2014 ~ exvessel_for_extrap * (cpi_2014/cpi_2012),
                                       year == 2015 ~ exvessel_for_extrap * (cpi_2015/cpi_2012),
                                       year == 2016 ~ exvessel_for_extrap * (cpi_2016/cpi_2012),
                                       year == 2017 ~ exvessel_for_extrap * (cpi_2017/cpi_2012),
                                       year == 2018 ~ exvessel_for_extrap * (cpi_2018/cpi_2012))) %>%
  dplyr::select(-exvessel, -exvessel_for_extrap) %>%
  rename(exvessel = exvessel_adjusted)

@@

# Look at NAs - 10,023 entries
combine_nas <- combine_catch_price_dat %>%
  dplyr::filter(is.na(exvessel_inclusive))

@@

# Calculate revenues and export
landed_value_dat <- combine_catch_price_dat %>%
  dplyr::select(-exvessel, -exvessel_group_avg, -exvessel_group_avg_allyears) %>%
  rename(exvessel = exvessel_inclusive) %>%
  mutate(landed_value = value*exvessel) %>%
  group_by(iso3, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(landed_value, na.rm = T)) %>%
  mutate(variable = "landed_value",
         units = paste(year, "US$"),
         source = "FAO (2020); Melnychuk et al. (2017)")

@@

## RAM Legacy Stock Assessment database

```{r}
# Load B/Bmsy timeseries from the RAM legacy stock assessment database
ram_dat_bv_bmsy <- read_csv(paste0(ram_sadb_path, "BvB.csv")) %>%
  rename(year = X1) %>%
  gather(stockid, bv_bmsy, -year) %>%
  na.omit() %>%
  group_by(stockid) %>%
  mutate(year_max = max(year)) %>%
  ungroup() %>%
  dplyr::filter(year == year_max)

# Load lookup table applying FAO regions to the RAM stocks
ram_lookup <- read_csv(fao_regions_for_ram_lookup_path)

# Apply to RAM data and create a new row for each FAO region
ram_dat_bv_bmsy_edit <- ram_dat_bv_bmsy %>%
  left_join(ram_lookup, by = "stockid") %>%
  mutate(region_fao = fao_area) %>%
  mutate(region_fao_split = str_split(region_fao, ";")) %>%
  unnest(cols = c(region_fao_split)) %>%
  group_by(stockid) %>%
  mutate(n_regions = n_distinct(region_fao_split)) %>%
  ungroup()

# Save clean
write_csv(ram_dat_bv_bmsy_edit, paste0(data_dir, "RAMLDB_v4.491_bvmsy_tidy.csv"))
```

## Estimates of global stock status from Costello et al. (2016)

```{r}
# Load data
costello_upsides_dat <- read_csv(costello_upsides_path) %>%
  clean_names(case = "snake")

# We need to split the entries that span multiple FAO regions 
costello_upsides_dat_edit <- costello_upsides_dat %>%
  mutate(region_fao_chr = as.character(region_fao),
         n_regions = nchar(region_fao_chr)/2,
         region_fao_split = str_split(gsub("(.{2})", "\\1 ", region_fao_chr), " ")) %>%
  unnest(cols = c(region_fao_split)) %>%
  mutate(region_fao_split = as.numeric(region_fao_split)) %>%
  dplyr::filter(region_fao_split != "") %>%
  mutate(iso3 = case_when(country == "Channel Islands" ~ "JEY",
                          country == "Micronesia" ~ "FSM",
                          country == "Zanzibar" ~ "TZA",
                          TRUE ~ countrycode(country, "country.name", "iso3c")))

@@
